---
title: "Principal Component Analysis (PCA)"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())

```

The following is PCA analysis on raw data retrieved from Bloomberg. 

```{r Import Library, include=FALSE}
rm(list=ls())
library(readxl)
library(corrr)
library(ggcorrplot)
library(ggplot2)
library(reshape2)
library(factoextra)


```

# Import data

```{r Import Data}
# input raw data
dataset = read.csv('../data/data.csv')
dataset = na.omit(dataset)        # handle NA values
remove.columns = c('date', 'daily_trading_value', 'UBB', 'LBB', 'BollMA', 'MA_5', 'MA_25', 'MACD', 'sig_9', 'diff', 'golden_cross', 'death_cross', 'sixth_day_return')
raw.data = dataset[ ,!(colnames(dataset) %in% remove.columns)]        # exclude date in dataset raw_data
head(raw.data)

# normalize raw data
raw.data.normalized <- scale(raw.data)

head(as.data.frame(raw.data.normalized))
```

**correlation heatmap on raw data & normalized raw data (same plot)**

```{r Raw Data Heatmap}
# plot correlation heatmap on raw data
raw.data.corr.matrix <- cor(raw.data)
ggcorrplot(raw.data.corr.matrix)
normalized.data.corr.matrix <- cor(raw.data.normalized)
ggcorrplot(normalized.data.corr.matrix)

```

```{r include=FALSE}
# plot correlation heatmap on normalized raw data
# raw.data.normalized.corr.matrix <- cor(raw.data.normalized)
# ggcorrplot(raw.data.normalized.corr.matrix)
# ggplot(melt(cor(raw.data)), aes(x = Var1, y = Var2, fill = value)) + geom_tile() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + scale_fill_gradient(low = "#56b1f7", high = "#16314b") 

```

## Method 1: PCA based on raw data diff

```{r Data Difference}
# get the daily changes of raw dataframe raw.data
raw.data.diff = as.data.frame(apply(raw.data, 2, diff))
head(raw.data.diff)

as.data.frame(apply(raw.data.diff, 2, mean))

```

**correlation heatmap on data difference**

```{r Data Difference Heatmap}
# heatmap of correlation on raw data
raw.data.diff.corr.matrix <- cor(raw.data.diff)
ggcorrplot(raw.data.diff.corr.matrix)

```


```{r}
# PCA analysis on raw.data.diff
out1 = princomp(raw.data.diff)

# visualization of PAC result
as.data.frame(out1$loadings[,1:4])


fviz_eig(out1, addlabels = TRUE)

# graph of the variables
fviz_pca_var(out1, col.var = "cos2", gradient.cols = c("black", "yellow", "red"), repel = TRUE)

# contribution of each variable
fviz_cos2(out1, choice = "var", axes = 1:2)
```

## Method 2: PCA analysis on correlation of data difference

```{r include=FALSE}
# PCA analysis on corr.matrix
out2 = princomp(raw.data.diff.corr.matrix)
summary(out2)

# visualization of PAC result
(PCA_componants <- as.data.frame(out2$loadings[,1:3]))
setwd('../data')
write.csv(PCA_componants, file = 'diff_corr_PCA_componants.csv')

fviz_eig(out2, addlabels = TRUE)

# graph of the variables
fviz_pca_var(out2, col.var = "cos2", gradient.cols = c("black", "yellow", "red"), repel = TRUE)

# contribution of each variable
fviz_cos2(out2, choice = "var", axes = 1:2)

```

## Method 3: PCA based on normalized raw data diff (Choice 1)

```{r Normalized Data Difference}
# get the daily changes of normalized dataframe 
normalized.data.diff = as.data.frame(apply(raw.data.normalized, 2, diff))
head(normalized.data.diff)

as.data.frame(apply(normalized.data.diff, 2, mean))

```

**correlation heatmap on normalized data difference** Same as correlation heatmap on data difference

```{r Normalized Data Difference Heatmap, include=FALSE}
# heatmap of correlation on normalized data
normalized.data.diff.corr.matrix <- cor(normalized.data.diff)
ggcorrplot(normalized.data.diff.corr.matrix)

```


```{r}
# PCA analysis on normalzied.data.diff
out3 = princomp(normalized.data.diff)
summary(out3)

# visualization of PAC result
(PCA_componants <- as.data.frame(out3$loadings[,1:2]))
setwd('../data')
write.csv(PCA_componants, file = 'normalized_data_diff_PCA_componants.csv')

fviz_eig(out3, addlabels = TRUE)

# graph of the variables
fviz_pca_var(out3, col.var = "cos2", gradient.cols = c("black", "yellow", "red"), repel = TRUE)

# contribution of each variable
fviz_cos2(out3, choice = "var", axes = 1:2)
```

## Method 4: PCA analysis on correlation of nromalized data difference

Same as PCA analysis on correlation of data difference.

```{r include=FALSE}
# PCA analysis on corr.matrix
out4 = princomp(normalized.data.diff.corr.matrix)

# visualization of PAC result
as.data.frame(out4$loadings[,1:4])

fviz_eig(out4, addlabels = TRUE)

# graph of the variables
fviz_pca_var(out4, col.var = "cos2", gradient.cols = c("black", "yellow", "red"), repel = TRUE)

# contribution of each variable
fviz_cos2(out4, choice = "var", axes = 1:2)
```

## Method 5: PCA based on normalized data (Choice 2)
```{r}
# PCA analysis on normalzied.data
out5 = princomp(raw.data.normalized)
summary(out5)

# visualization of PAC result
(PCA_componants <- as.data.frame(out5$loadings[, 1:5]))
setwd('../data')
write.csv(PCA_componants, file = 'normalized_data_PCA_componants.csv')


# visualization of PAC result
as.data.frame(out5$loadings[,1:4])

fviz_eig(out5, addlabels = TRUE)

# graph of the variables
fviz_pca_var(out5, col.var = "cos2", gradient.cols = c("black", "yellow", "red"), repel = TRUE)

# contribution of each variable
fviz_cos2(out3, choice = "var", axes = 1:2)
```

## Method 6: PCA based on correlation of normalized data
```{r, include=FALSE}
normalized.data.corr.matrix <- cor(raw.data.normalized)

# PCA analysis on normalzied.data
out6 = princomp(normalized.data.corr.matrix)
summary(out6)

# visualization of PAC result
(PCA_componants <- as.data.frame(out6$loadings[,1:4]))
setwd('../data')
write.csv(PCA_componants, file = 'normalized_data_corr_PCA_componants.csv')


# visualization of PAC result
as.data.frame(out6$loadings[,1:4])

fviz_eig(out6, addlabels = TRUE)

# graph of the variables
fviz_pca_var(out6, col.var = "cos2", gradient.cols = c("black", "yellow", "red"), repel = TRUE)

# contribution of each variable
fviz_cos2(out3, choice = "var", axes = 1:2)
```

## References

[Principal Component Analysis in R Tutorial](https://www.datacamp.com/tutorial/pca-analysis-r)  
[Principal Component Analysis (PCA) 101, using R](https://towardsdatascience.com/principal-component-analysis-pca-101-using-r-361f4c53a9ff)
