---
title: "Principal Component Analysis (PCA)"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())

```

The following is PCA analysis on raw data retrieved from Bloomberg.

```{r Import Library, include=FALSE}
rm(list=ls())
library(readxl)
library(corrr)
library(ggcorrplot)
library(ggplot2)
library(reshape2)
library(factoextra)

```

# Import data

```{r Import Data}
# input raw data
dataset = read_xlsx('../data/raw_data_adjusted.xlsx')
dataset = na.omit(dataset)        # handle NA values
raw.data = dataset[ ,!(colnames(dataset) %in% c('date'))]        # exclude date in dataset raw_data
head(raw.data)

# normalize raw data
raw.data.normalized <- scale(raw.data)
head(as.data.frame(raw.data.normalized))
```

**correlation heatmap on raw data & normalized raw data (same plot)**

```{r Raw Data Heatmap}
# plot correlation heatmap on raw data
raw.data.corr.matrix <- cor(raw.data)
ggcorrplot(raw.data.corr.matrix)

```

```{r include=FALSE}
# plot correlation heatmap on normalized raw data
# raw.data.normalized.corr.matrix <- cor(raw.data.normalized)
# ggcorrplot(raw.data.normalized.corr.matrix)
# ggplot(melt(cor(raw.data)), aes(x = Var1, y = Var2, fill = value)) + geom_tile() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + scale_fill_gradient(low = "#56b1f7", high = "#16314b") 

```

## Method 1: PCA based on raw data

```{r Data Difference}
# get the daily changes of raw dataframe raw.data
raw.data.diff = as.data.frame(apply(raw.data, 2, diff))
head(raw.data.diff)

as.data.frame(apply(raw.data.diff, 2, mean))

```

**correlation heatmap on data difference**

```{r Data Difference Heatmap}
# heatmap of correlation on raw data
raw.data.diff.corr.matrix <- cor(raw.data.diff)
ggcorrplot(raw.data.diff.corr.matrix)

```

### PCA analysis on data difference

```{r}
# PCA analysis on raw.data.diff
out1 = princomp(raw.data.diff)

# visualization of PAC result
as.data.frame(out1$loadings[,1:4])


fviz_eig(out1, addlabels = TRUE)

# graph of the variables
fviz_pca_var(out1, col.var = "cos2", gradient.cols = c("black", "yellow", "red"), repel = TRUE)

# contribution of each variable
fviz_cos2(out1, choice = "var", axes = 1:2)
```

### *!!!!PCA analysis on correlation of data difference*

```{r}
# PCA analysis on corr.matrix
out2 = princomp(raw.data.diff.corr.matrix)

# visualization of PAC result
as.data.frame(out2$loadings[,1:4])

fviz_eig(out2, addlabels = TRUE)

# graph of the variables
fviz_pca_var(out2, col.var = "cos2", gradient.cols = c("black", "yellow", "red"), repel = TRUE)

# contribution of each variable
fviz_cos2(out2, choice = "var", axes = 1:2)
```

## Method 2: PCA based on normalized raw data

```{r Normalized Data Difference}
# get the daily changes of nromalized dataframe 
normalized.data.diff = as.data.frame(apply(raw.data.normalized, 2, diff))
head(normalized.data.diff)


as.data.frame(apply(normalized.data.diff, 2, mean))

```

**correlation heatmap on normalized data difference** Same as correlation heatmap on data difference

```{r Normalized Data Difference Heatmap, include=FALSE}
# heatmap of correlation on normalized data
normalized.data.diff.corr.matrix <- cor(normalized.data.diff)
ggcorrplot(normalized.data.diff.corr.matrix)

```

### PCA analysis on normalized data difference

```{r}
# PCA analysis on normalzied.data.diff
out3 = princomp(normalized.data.diff)

# visualization of PAC result
as.data.frame(out3$loadings[,1:4])

fviz_eig(out3, addlabels = TRUE)

# graph of the variables
fviz_pca_var(out3, col.var = "cos2", gradient.cols = c("black", "yellow", "red"), repel = TRUE)

# contribution of each variable
fviz_cos2(out3, choice = "var", axes = 1:2)
```

### PCA analysis on correlation of nromalized data difference

Same as PCA analysis on correlation of data difference.

```{r include=FALSE}
# PCA analysis on corr.matrix
out4 = princomp(normalized.data.diff.corr.matrix)

# visualization of PAC result
as.data.frame(out4$loadings[,1:4])

fviz_eig(out4, addlabels = TRUE)

# graph of the variables
fviz_pca_var(out4, col.var = "cos2", gradient.cols = c("black", "yellow", "red"), repel = TRUE)

# contribution of each variable
fviz_cos2(out4, choice = "var", axes = 1:2)
```

## References

<https://www.datacamp.com/tutorial/pca-analysis-r>
